\---
title: DS 6372 - Project 1
author: "Scott & Holmes"
output: word_document
date: "9/26/22"
---

__Libraries & Setup__
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(tidyverse)
library(stringr)
library(corrplot)
library(leaps)

library(httpgd)
hgd()
hgd_browse()
```


__Data Entry & Handeling__
```{r}
# reading in the csvs
msrp <- read.csv(file.choose())
# msrp <- read.csv("./Data Files//msrp.csv") #read csv

msrp$Market_Category <-
    str_replace_all(
        msrp$Market.Category,
        c(
            ".*Luxury.*" = "Luxury",
            ".*Crossover.*" = "Crossover",
            ".*High-Performance.*" = "High-Performance",
            ".*Performance.*" = "Performance",
            ".*Exotic.*" = "Exotic",
            ".*Hybrid.*" = "Hybrid",
            ".*Hatchback.*" = "Hatchback",
            ".*Factory Tuner.*" = "Factory Tuner",
            ".*Diesel.*" = "No Category",
            "N/A" = "No Category"
        )
    )

# Creating categorical variables
msrp <-
    msrp %>%
    mutate(
        Luxury = case_when(Market_Category == "Luxury" ~ 1, TRUE ~ 0),
        Crossover = case_when(Market_Category == "Crossover" ~ 1, TRUE ~ 0),
        High_Performance = case_when(Market_Category == "High-Performance" ~ 1, TRUE ~ 0),
        Performance = case_when(Market_Category == "Performance" ~ 1, TRUE ~ 0),
        Exotic = case_when(Market_Category == "Exotic" ~ 1, TRUE ~ 0),
        Hybrid = case_when(Market_Category == "Hybrid" ~ 1, TRUE ~ 0),
        Factory_Tuner = case_when(Market_Category == "Factory Tuner" ~ 1, TRUE ~ 0)
    )

# Cleaning up Fuel types to be larger categories
msrp$Fuel_Type <-
    str_replace_all(
        msrp$Engine.Fuel.Type,
        c(
            ".*premium unleaded .*required.*" = "Premium Required",
            ".*regular.*" = "Regular",
            ".*premium unleaded .*recommended.*" = "Premium Recommended",
            ".*diesel.*" = "Diesel",
            ".*natural gas.*" = "Natural Gas",
            "electric.*" = "Electric",
            "flex-fuel.*" = "Flex-fuel"
        )
    )

msrp <-
    msrp %>%
    mutate(Fuel_Type = case_when(Model == "Verona" ~ "Regular", TRUE ~ Fuel_Type))


# pulling from the bureau of labor statistics to adjust pricing based on year/new car estimates
# (source: https://data.bls.gov/pdq/SurveyOutputServlet)

# car_inflation <-
#     data.frame(
#         Year = c(
#             1990, 1991, 1992, 1993, 1994, 1995, 1996, 1997, 1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005,
#             2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017
#         ),
#         inflation = c(
#             1.210809308, 1.142925336, 1.120985403, 1.096947855, 1.06400961, 1.041105909, 1.022624209,
#             1.018655745, 1.024154902, 1.028123365, 1.028860365, 1.033225675, 1.047568834, 1.062195456, 1.067581227,
#             1.06179861, 1.06400961, 1.073050337, 1.087069783, 1.07734308, 1.061137577, 1.034755801, 1.018777634,
#             1.008224356, 1.004880643, 0.99902886, 0.997515175, 1
#         )
#     )
# msrp <- merge(x = msrp, y = car_inflation, by = "Year")
# msrp$Adjusted_MSRP <- msrp$inflation * msrp$MSRP

msrp$adjusted.MSRP <- as.factor(msrp$Year)
msrp$Make <- as.factor(msrp$Make)
msrp$Model <- as.factor(msrp$Model)
msrp$Engine.Cylinders <- as.factor(msrp$Engine.Cylinders)
msrp$Transmission.Type <- as.factor(msrp$Transmission.Type)
msrp$Driven_Wheels <- as.factor(msrp$Driven_Wheels)
msrp$Number.of.Doors <- as.factor(msrp$Number.of.Doors)
msrp$Vehicle.Size <- as.factor(msrp$Vehicle.Size)
msrp$Vehicle.Style <- as.factor(msrp$Vehicle.Style)
msrp$Market_Category <- as.factor(msrp$Market_Category)
msrp$Fuel_Type <- as.factor(msrp$Fuel_Type)

msrp$Luxury <- as.factor(msrp$Luxury)
msrp$Crossover <- as.factor(msrp$Crossover)
msrp$High_Performance <- as.factor(msrp$High_Performance)
msrp$Performance <- as.factor(msrp$Performance)
msrp$Exotic <- as.factor(msrp$Exotic)
msrp$Hybrid <- as.factor(msrp$Hybrid)
msrp$Factory_Tuner <- as.factor(msrp$Factory_Tuner)

# Replacing Engine Cylinder for Electric vehicles with NULL
# msrp$Engine.Cylinders <- msrp$Engine.Cylinders %>% replace_na()
# msrp["Engine.Cylinders"][msrp["Engine.Cylinders"] == 0] <- "Not Available"
```


__EDA__
```{r}
msrp %>%
    ggplot(aes(x = Year, y = MSRP)) +
    geom_point() +
    theme_minimal() +
    ggtitle("Scatterplot - MSRP by Year")
```

There appears to be a significant difference in the way data with gathered before and after the year 2000. For this reason, we have decided to focus our model on years 2001 and on. We believe this provides the most use to the general audience because it aligns to the most recent years to our current time period. 

Furthermore, we can see a very limited sample size for cars with an MSRP greater than $500,000. We have decided to focus our model on cars below this threshold.

```{r}
# Filtering dataset and adding log variables
msrp_trim <-
    msrp %>%
    filter(Year > 2000) %>%
    filter(MSRP < 500000) %>%
    mutate(log_msrp = log(MSRP)) %>%
    select(-c(Engine.Fuel.Type, Market.Category))
str(msrp_trim)

# N/A value totals
# from https://stackoverflow.com/questions/24027605/determine-the-number-of-na-values-in-a-column
msrp_trim %>%
    map_df(function(x) sum(is.na(x))) %>%
    gather(feature, num_nulls) %>%
    print(n = 100) %>%
    filter(num_nulls > 0)

msrp_trim %>% filter(is.na(Engine.HP)) %>% group_by(Fuel_Type) %>% summarise(count = n())
msrp_trim %>% filter(is.na(Engine.Cylinders)) %>% group_by(Fuel_Type) %>% summarise(count = n())
msrp_trim %>% filter(is.na(Number.of.Doors)) %>% group_by(Fuel_Type) %>% summarise(count = n())
## Electirc cars are contributing to the majority of N/A values

msrp_trim %>%
    mutate(electric = ifelse(Fuel_Type == "Electric", TRUE, FALSE)) %>%
    filter(electric == TRUE) %>%
    ggplot(aes(x = Year, fill = electric)) + geom_bar() +
    theme_minimal()
    ggtitle("Electric Cars by Years") +
    ylab("Count")
## Electric cars represent a small portion of our overall sample. They are also only relevant years 2014-2017

msrp_trim2 <-
    msrp_trim %>%
    filter(Fuel_Type != "Electric")

msrp_trim2 %>%
    map_df(function(x) sum(is.na(x))) %>%
    gather(feature, num_nulls) %>%
    print(n = 100) %>%
    filter(num_nulls > 0)

view(msrp_trim2 %>% filter(is.na(Engine.HP))) # Action - Remove NA rows
view(msrp_trim2 %>% filter(is.na(Engine.Cylinders))) # Action - Replace NA with correct value for car
view(msrp_trim2 %>% filter(is.na(Number.of.Doors))) # Action - Replace NA with correct value for car

# RX-8 from all Years had 4-cylinder engines.
msrp_trim2$Engine.Cylinders <- msrp_trim2$Engine.Cylinders %>% replace_na("4")
# Ferrari FF from 2013 had 2 doors
msrp_trim2$Number.of.Doors <- msrp_trim2$Number.of.Doors %>% replace_na("2")


# Engine HP is the most highly correlated continuous variable with MSRP. We dropped all N/A rows.
msrp_trim3 <-
    msrp_trim2 %>%
    filter(!is.na(Engine.HP))

# Make
msrp_trim3 %>%
    ggplot(aes(x = Make, y = log_msrp)) +
    geom_boxplot()

msrp_by_make <-
    msrp_trim %>%
    group_by(Make) %>%
    summarise(Mean = mean(log_msrp))

# Creating new Make groups by mean msrp
msrp_by_make <-
    msrp_by_make %>%
    mutate(
        make_group =
            ifelse(Mean < 10, "<10",
                ifelse(Mean < 11, "<11",
                    ifelse(Mean < 12, "<12", ">12")
                )
            )
    ) %>%
    select(c(Make, make_group))

# Merging new Make Group into main dataset
msrp_trim4 <- merge(x = msrp_trim3, y = msrp_by_make, by = "Make", all.x = TRUE)
view(msrp_trim4)

msrp_trim4$make_group <- as.factor(msrp_trim4$make_group)
```

After grouping the makes by their mean adjusted log(msrp), we can see a very clear correlation between make group and adjusted log(msrp).

__Test, Train, Validation Splits__
```{r}
# Setup
set.seed(200)
msrp_train <-
    msrp_trim3 %>%
        select(c(Year, Engine.HP, Engine.Cylinders, Transmission.Type, Driven_Wheels, Number.of.Doors, Vehicle.Size,
            Vehicle.Style, Popularity, Market_Category, Fuel_Type, log_adj_msrp, make_group))

# Dataset
index <- sample(seq_len(dim(msrp_train)[1]), dim(msrp_train)[1] * .8, replace = FALSE)
train <- msrp_train[index, ]
remainder <- msrp_train[-index, ]

index2 <- sample(seq_len(dim(remainder)[1]), dim(remainder)[1] * .5, replace = FALSE)
test <- remainder[index2, ]
validate <- remainder[-index2, ]

dim(train)
dim(test)
dim(validate)
```



__EDA Continued__
```{r}
# MSRP Review
msrp_trim3 %>%
    ggplot(aes(x = MSRP)) +
    geom_histogram() +
    theme_minimal() +
    ggtitle("Histogram - MSRP") +
    ylab("Count")

msrp_trim %>%
    ggplot(aes(x = log_msrp)) +
    geom_histogram() +
    theme_minimal() +
    ggtitle("Histogram - MSRP") +
    xlab("log(MSRP)") +
    ylab("Count")

# Significantly more cars from 2015-2017 than other years.
msrp_trim3 %>%
    group_by(Year) %>%
    summarise(count = n())
```

Even after filtering to Years > 2000 and MSRP < $500,000, MSRP is still heavily right skewed. We decided to introduce a log transformation to normalize MSRP distirbution. After the transformation, the distribution is still slright right skewed. We believe that with the sample size, the properties of the Critical Limit Theorem will account for any remaining skewness in our data. 

```{r}
# Continuous Variables
num_cols <-
    msrp_trim %>%
    select(where(is.numeric)) %>%
    select(-c(MSRP, Adjusted_MSRP, inflation))
str(num_cols)
# Categorical Variables
cat_cols <-
    msrp_trim %>%
    select_if(negate(is.numeric))
str(cat_cols)
```

There are some concerns of collinearity between Engine HP and both Highway and City MGP. There is also significant coorelation between Highway and City MPG. Both have an equal correlation with our response variable. We have decided to remove City MPG from the model to account for it's collinearity with Engine HP and Highway MPG.

```{r}
# Testing for multicollinearity in the continuous variables
cor_matrix <- cor(num_cols, use = "complete.obs")
corrplot.mixed(cor_matrix)
corrplot(cor_matrix, method = "ellipse", type = "upper")
pairs(num_cols)

msrp_trim %>% ggplot(aes(x = Engine.HP, y = log_msrp)) + geom_point()

```

```{r}
# Exploring Categorical Variables
## Fuel Type
msrp_trim %>% ggplot(aes(x = Engine.HP, y = log_msrp, color = Fuel_Type)) +
    geom_point() +
    geom_smooth(method = "lm")

msrp_trim %>%
    group_by(Fuel_Type) %>%
    summarise(count = n())
### Concerns about limited Natrual Gas and Electric Vehicle Fuel Types
msrp_trim %>% ggplot(aes(x = Popularity, y = log_adj_msrp)) +
    geom_point()

msrp_trim %>% ggplot(aes(x = Year, y = log_adj_msrp)) +
    geom_point()
```

Multicollinearity
 - Both MPG variables and Engine HP
 - City MPG and Highway MPG
    - Remove both MPG variables from the model in favor of Engine HP

 - Remove lines with Adjusted_MSRP > $500,000
 - Remove lines with no Horsepower recorded
 - If Fuel_Type is needed, remove Natural Gas
 - If removing Horesepower = 0 





Model has way too many levels and would immediately create an issue of overfitting if we included it in the model.

```{r}
# Cylinders
msrp_trim2 %>% ggplot(aes(x = Engine.Cylinders, y = log_adj_msrp)) +
    geom_boxplot()
summary(lm(log_adj_msrp ~ Engine.Cylinders, data = msrp_trim2))

# Market_Category
msrp_trim2 %>% ggplot(aes(x = Market_Category, y = log_adj_msrp)) +
    geom_boxplot()
summary(lm(log_adj_msrp ~ Market_Category, data = msrp_trim2))

# Transmission Type
msrp_trim2 %>% ggplot(aes(x = Transmission.Type, y = log_adj_msrp)) +
    geom_boxplot()
summary(lm(log_adj_msrp ~ Transmission.Type, data = msrp_trim2))

# Driven_Wheels
msrp_trim2 %>% ggplot(aes(x = Driven_Wheels, y = log_adj_msrp)) +
    geom_boxplot()
summary(lm(log_adj_msrp ~ Driven_Wheels, data = msrp_trim2))

# Vehicle.Size
msrp_trim2 %>% ggplot(aes(x = Vehicle.Size, y = log_adj_msrp)) +
    geom_boxplot()
summary(lm(log_adj_msrp ~ Vehicle.Size, data = msrp_trim2))

# Number.of.Doors
msrp_trim2 %>% ggplot(aes(x = Number.of.Doors, y = log_adj_msrp)) +
    geom_boxplot()
summary(lm(log_adj_msrp ~ Number.of.Doors, data = msrp_trim2))

# Categorical Statistics
t(aggregate(log_adj_msrp ~ Engine.Fuel.Type, data = msrp_trim, summary))

t(aggregate(log_adj_msrp ~ Market.Category, data = msrp_trim, mean))

msrp_trim %>%
    count(Engine.Cylinders, sort = TRUE)

# Charts
msrp_trim %>%
    ggplot(aes(x = Year, y = log_adj_msrp, color = Market_Category)) +
    geom_point()

msrp_trim %>%
    ggplot(aes(x = log_adj_msrp)) +
    geom_histogram()

msrp_train %>% ggplot(aes(x = Engine.HP, y = log_adj_msrp, color = Fuel_Type)) + geom_point()
```

__Model Building__
```{r}
# Feature Selection
model <- lm(log_adj_msrp ~ ., data = train)
summary(model)
reg_fwd <- regsubsets(log_adj_msrp ~ .,
    data = train, method = "forward", nvmax = 100)
summary(reg_fwd)$adjr2
summary(reg_fwd)$rss
summary(reg_fwd)$bic

par(mfrow = c(1, 3))
bics <- summary(reg_fwd)$bic
plot(1:51, bics, type = "l", ylab = "BIC", xlab = "# of predictors")
index <- which(bics == min(bics))
points(index, bics[index], col = "red", pch = 10)

adjr2 <- summary(reg_fwd)$adjr2
plot(1:51, adjr2, type = "l", ylab = "Adjusted R-squared", xlab = "# of predictors")
index <- which(adjr2 == max(adjr2))
points(index, adjr2[index], col = "red", pch = 10)

rss <- summary(reg_fwd)$rss
plot(1:51, rss, type = "l", ylab = "train RSS", xlab = "# of predictors")
index <- which(rss == min(rss))
points(index, rss[index], col = "red", pch = 10)

# Prediction Function from Turner
predict.regsubsets <- function(object, newdata, id, ...) {
    form <- as.formula(object$call[[2]])
    mat <- model.matrix(form, newdata)
    coefi <- coef(object, id = id)
    xvars <- names(coefi)
    mat[, xvars] %*% coefi
}

testASE <- c()
# note my index is to 20 since that what I set it in regsubsets
for (i in 1:51) {
    predictions <- predict.regsubsets(object = reg_fwd, newdata = test, id = i)
    testASE[i] <- mean((test$log_adj_msrp) - predictions)^2
}
par(mfrow = c(1, 1))
plot(1:51, testASE, type = "l", xlab = "# of predictors", ylab = "test vs train ASE", ylim = c(0, 1))
index <- which(testASE == min(testASE))
points(index, testASE[index], col = "red", pch = 10)
rss <- summary(reg_fwd)$rss
lines(1:51, rss / dim(test)[1], lty = 3, col = "blue") # Dividing - ASE=RSS/sample size


reg.final <- regsubsets(log_adj_msrp ~ ., data = train, method = "forward", nvmax = 7)
coef(reg.final, 2)
final.model <- lm(log_adj_msrp ~ Engine.HP + Market_Category, data = msrp_trim3)
summary(final.model)

plot(exp(final.model$fitted.values),exp(msrp_trim3$log_adj_msrp),xlab="Predicted",ylab="adj_msrp",xlim=c(0,400000),ylim=c(0,400000))
lines(c(0, 400000), c(0, 400000), col = "red")
```
